Estimating dark matter halo concentration using the integrated mass profile

### 2015-12-17 paper/paper.tex
nueva version revisada del texto

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{error.pdf}
\end{center}
\caption{Average value of the relative error in the concentration
  estimate, $\avg{|D|}$, as a function of the particle number $N$ in
  the set of mock halos. Different symbols represent different
  methods. Our method provide the most accurate estimate at fixed
  particle number $N$.  {\it Con respecto a esta figura tengo una
    pregunta. Yo ajusté los perfiles con menos de 200 partulas? no
    recuerdo!}
    \label{fig:error}}
\end{figure}

\label{sec:data}
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.32\textwidth]{conc_mass_vel.pdf}
    \includegraphics[width=0.32\textwidth]{conc_mass_dens.pdf}
    \includegraphics[width=0.32\textwidth]{conc_dens_vel.pdf}
  \end{center}
  \caption{Comparison between the concentrations measured by our
    method and the maximum velocity (left) and density (middle)
    methods. The line indicates the equal value between the two
    techniques. The right panel compares the results of the maximum
    velocity and density methods.
  \label{fig:mdv}}
\end{figure*}

@@

Figure \ref{fig:concentration} shows the results of the
mass-concentration relationship.

@@

%{\em Aqui hay algo importante que hay que entender antes de discutir
%  esta gráfica , y es la razón por la cual el metodo del perfil de
%  masa da valores de concentración tan bajos a baja masa y comparables
%  a los de los otros dos métodos a alta masa. Es importante entender
%  la razón de esto!.  Se me ocurren unas preguntas para tratar de
%  entender. \\ 1) La grafica de c-M \ref{fig:concentration} usa todos
%  los halos o solo los halos relajados? La inclusión de halos no
%  relajados introduce in bias en la mediana en Cvir, así que vale la
%  pena inclusive agregar una linea mas a la gráfica con los halos
%  relajados y sin ellos. Para los otros dos métodos ya se sabe que
%  hace la población de halos no relajados en la medida así que no creo
%  necesario agregarlas para esos métodos. \\ Hay que explicar un poco
%  mas la gráfica en el sentido de que se grafica y como se
%  calculó. mean concentration?, Median concentrations? o mean in
%  logscale?  Mas bins de cuanto? Mínimo cuantos halos por mass bin?,
%  Halos relajados en todos los casos, o indistintos?}

@@

The first benchmark was performed on mock halos generated with known
concentration values for different particle numbers.  For all methods,
the accuracy in retrieveing the input concentration increases with the
number of particles as summarized in Figure \ref{fig:error}.  Our
method systematically shows smaller errors in the retrieved values
compared to the other two methods, going from $10\%$ for halos with
$20$ particles down to $0.01\%$ for halos with $2\times 10^4$
particles.  In our method the average error on the retrieved values
decreases with the number of particles as $\langle|D|\rangle\propto
N^{-1}$.  The error in density method initially behaves as
$\langle|D|\rangle\propto N^{-3/2}$ but at $N=2\times 10^3$ it
saturates to $0.1\%$.  For the velocity method the error behaves
closely to $\langle|D|\rangle\propto N^{-1/2}$.  We also gauged the
effect of different concentration parameters at fixed particle
numbers.  We found larger offsets for larger input concentrations.
The offset goes in the sense that the measured concentration tends to
be underestimated.  Nevertheless, the dominant effect on the
measurement is the particle number.

We used a N-body cosmological simulation to test the impact of our
method on the mass-concentration relationship.  Although the three
methods are in general agreement there are noticeable differences in
the mass concentration relationship.

>>>>

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{error.pdf}
\end{center}
\caption{Average value of the relative error in the concentration
  estimate, $\avg{|D|}$, as a function of the particle number $N$ in
  the set of mock halos. Different symbols represent different
  methods. Our method provide the most accurate estimate at fixed
  particle number $N$.  
    \label{fig:error}}
\end{figure}

\label{sec:data}
\begin{figure*}
  \begin{center}
    \includegraphics[width=0.32\textwidth]{conc_mass_vel.pdf}
    \includegraphics[width=0.32\textwidth]{conc_mass_dens.pdf}
    \includegraphics[width=0.32\textwidth]{conc_dens_vel.pdf}
  \end{center}
  \caption{Comparison between the concentrations measured by our
    method and the maximum velocity (left) and density (middle)
    methods. The solid line indicates the equal value between the two
    techniques. The dashed line shows the upper 15 per cent difference
    between two methods.
  \label{fig:comparison}}
\end{figure*}

@@

Figure \ref{fig:comparison} shows the comparison between the
concentration values obtained by the three different methods. 
The diagonal line marks a one-to-one correspondence if all the values
were equal.
The three methods show a broad general agreement. 
The largest difference is present at high concentration $c>5$ where
our method produces systematic lower values than the other two
methods. Most of the differences are at the $20\%$ level, with some
extreme cases for the highest concentration values with $50\%$
differences.  The velocity and density methods produce consistent
values within a $20\%$ scatter, only some halos with very high
concentrations $c\sim 10$ have a $15\%$ lower concentration with the
velocity method.

We explored halo shape and relaxation as a source of this
differences. We found shape does not bias the concentration
estimation; relaxation, quantified by the
offset between the center of mass and minimum of the potential, can
only explain at most $5\%$ of the offset between methods.  We note
that the highest concentration values correspond to the smallest halos
with $\sim 200$ particles, a regime where both the density and
velocity methods are reliable only the $5\%$ in a perfect spherical
halo.  From this we suggest that asserting with confidence the
systematic difference between the different methods should be
confirmed  a higher resolution simulation and larger statistics. We
leave this comparison for future work.

Figure \ref{fig:concentration} summarizes the previous results in the
mass-concentration relationship.

@@

@@

The first benchmark was performed on mock halos generated with known
concentration values for different particle numbers.  For all methods,
the accuracy in retrieveing the input concentration increases with the
number of particles as summarized in Figure \ref{fig:error}.  Our
method systematically shows smaller errors in the retrieved values
compared to the other two methods.

We used a N-body cosmological simulation to test the impact of our
method on the mass-concentration relationship.  Although the three
methods are in general agreement there are some noticeable differences
in the mass concentration relationship.
